---
title: "Updates April 11"
author: "Owen G. Ward"
date: "4/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
source(here("Experiments/utils.R"))
```

# Poisson Initialization

## Dense Setting

This is a quite easy simulation setting but the initialization procedure 
still leads to a slight improvement.
Here `no_init` corresponds to the random initialization, which shows a lot 
more variance.

```{r dense-sims, echo=FALSE, message=FALSE}
K <- 2
true_Mu <- matrix(0.05, 
                    nrow = K, ncol = K, byrow = T)
  diag(true_Mu) <- c(rep(0.5, K-1), 1)

print("Rate Matrix")
print(true_Mu)

exp_12_dense <- list.files(path = here("Experiments/exp_results/"), 
                           pattern = "exp_12_n0_dense")

exp_12_dense %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  group_by(init, n0) %>% 
  summarise(mean_ARI = mean(ARI), sd_ARI = sd(ARI))

exp_12_dense %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>%
  mutate(n0 = n0/2) %>% 
  ggplot(aes(as.factor(n0), ARI, fill = init)) +
  geom_boxplot() +
  geom_point(alpha = 0.25) +
  facet_wrap(vars(nodes, init)) +
  labs(x = "% training", fill = "Method") +
  scale_x_discrete(breaks = c(5, 10, 15, 20, 25))
```

We also consider a setting with $K=3$ which is harder for the random
starting point to recover the communities.

```{r k=3, echo=TRUE}
K <- 3
true_Mu <- matrix(0.05, 
                    nrow = K, ncol = K, byrow = T)
diag(true_Mu) <- 0.5:K + 1.2
print(true_Mu)  

exp_12_sim_2 <- list.files(path = here("Experiments/exp_results/"), 
                           pattern = "exp_12_n0_dense_apr_13")

exp_12_sim_2 %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  group_by(n0) %>% 
  summarise(mean(ARI), sd(ARI))

exp_12_sim_2 %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  ggplot(aes(as.factor(n0), ARI)) +
  geom_boxplot()

```


## Sparse Setting

Here $\rho=0.1,0.05,0.02$ and  `no_init` corresponds to the random initialization, which shows a lot 
more variance and doesn't recover the clusters as well.

Here for the small sparse network our initialization
procedure still struggles but as the number of nodes increases 
we can reliably recover the network structure well.

```{r sparse-plots, echo=FALSE, fig.height=12, fig.width=8}
true_Mu <- matrix(c(2, 0.05, 0.4, 1.5), 
                    nrow = K, ncol = K, byrow = T)
print("Rate Matrix")
print(true_Mu)

exp_12_sparse <- list.files(path = here("Experiments/exp_results/"), 
                           pattern = "exp_12_april_12")

exp_12_sparse %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  group_by(init, n0) %>% 
  summarise(mean_ARI = mean(ARI), sd_ARI = sd(ARI))

exp_12_sparse %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  group_by(init, nodes, n0, m0) %>% 
  # summarise(mean_ARI = mean(ARI), sd(ARI), median(ARI), n()) %>% 
  filter(nodes == 100) %>% 
  mutate(n0 = n0/2) %>% 
  mutate(n0 = as.factor(n0)) %>% 
  ggplot(aes(n0, ARI, fill = init)) +
  # geom_point(aes()) +
  geom_boxplot() +
  facet_wrap(vars(m0, sparsity), scales = "free") +
  labs(x = "% time used for Initialization",
       y = "Mean ARI",
       fill = "Scheme",
       title = "Total Nodes = 100",
       subtitle = "Top Number is number of nodes used in Init,
       Bottom is Level of Sparsity") +
  scale_x_discrete(breaks = c(5, 10, 15, 20, 25)) +
  ylim(c(-0.1, 1))


exp_12_sparse %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  group_by(init, nodes, n0, m0) %>% 
  # summarise(mean_ARI = mean(ARI), sd(ARI), median(ARI), n()) %>% 
  filter(nodes == 400) %>% 
  mutate(n0 = n0/2) %>% 
  mutate(n0 = as.factor(n0)) %>% 
  ggplot(aes(n0, ARI, fill = init)) +
  # geom_point(aes()) +
  geom_boxplot() +
  facet_wrap(vars(m0, sparsity), scales = "free") +
  labs(x = "% time used for Initialization",
       y = "Mean ARI",
       fill = "Scheme",
       title = "Total Nodes = 400",
       subtitle = "Top Number is number of nodes used in Init,
       Bottom is Level of Sparsity") +
  scale_x_discrete(breaks = c(5, 10, 15, 20, 25)) +
  ylim(c(-0.1, 1))

```

# Experiment for Figure 1 

```{r fig1, echo=TRUE}
fig1 <- list.files(path = here("Experiments/exp_results/"), 
                           pattern = "fig_1_exp_1_apr_13")

fig1 %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  mutate(window_size = as.factor(window_size)) %>% 
  drop_na(ARI) %>%
  group_by(Method) %>% 
  summarise(mean_ari = mean(ARI), sd_ari = sd(ARI))

fig1 %>% 
  map_dfr(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  mutate(window_size = as.factor(window_size)) %>% 
  drop_na(ARI) %>% 
  ggplot(aes(window_size, ARI, fill = Method)) +
  geom_boxplot() +
  facet_wrap(~Method, scales = "free") +
  ylim(c(-0.1, 1)) +
  labs(x = "Aggregation Window Size",
       title = "Aggregation Unable to Recover Communities")
```
