---
title: "Simulation Results"
author: "Owen G. Ward"
date: "11/8/2021"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(latex2exp)
source(here("Experiments/utils.R"))
```

Included are some initial simulation results. For all experimental
results shown below, a two class model was considered with a
homogeneous Poisson process, with rate matrix given by

$$
B = 
\begin{pmatrix}
2 & 0.05 \\
0.05 & 1
\end{pmatrix}
$$

and equal group proportions $(0.5, 0.5)$.

Here we discuss

- Recovery of the rate matrix under random initialization
- Convergence of the windowed ELBO
- Simulation results investigating the agreement between theoretical results
on the regret rate and observed regret.
- Performance of our algorithm in terms of online loss of predicted events.


# Recovery for random initialization

Although our method, along with many variational optimization methods,
relies 
heavily on the initialization, for the simulation scenario 
described above we see that good recoevy of $B$ can be achieved
in many scenarios,


```{r load-data}
exp_1_files <- list.files(path = here("Experiments/exp_results/"),
                          pattern = "exp1_time")

elbo_dat <- exp_1_files %>% 
  map(~readRDS(here("Experiments/exp_results/", .x))) %>%
  map(~.x %>%  tibble(
    clust = map_dbl(., "clust"),
    elbo = map(., "est_elbo"),
    time = map_dbl(., "time")
  ) %>% 
    mutate(sim = row_number()) %>% 
    select(sim, clust, time, elbo) %>% 
    unnest(cols = c(elbo)) %>% 
    group_by(sim) %>% 
    mutate(wind = 1:unique(time))) %>% 
  bind_rows()

```

## Convergence of the Window ELBO

The definition of the window ELBO here is given by

$$
ELBO(wind_m) = \frac{1}{events_{cum}}
\left[
\sum_{(i,j) \in A} \sum_{k,l} \tau_{ik} \tau_{jl} \bigg \{\int_{dT m}^{dT(m+1)} \log \lambda_{kl}(t) dN_{ij}(t) 
  - \int_{dT m}^{dT(m+1)} \lambda_{kl} (t) dt \bigg \} \\
  + \sum_{i=1}^{n} \sum_{k=1}^{K} \tau_{ik} \log{\pi_k / \tau_{ik}}
  \right],
$$

with $events_{cum}$ the number of events observed up to $dT(m+1)$.
Although this is slightly different to what we define as the
window ELBO in the paper, this is what we have always used in all
experiments previously, as opposed to $|A|$.

```{r plot-elbo}
elbo_dat %>% 
  filter(time == 50) %>% 
  mutate(correct = ifelse(clust == 1, "Recover", "Don't")) %>% 
  ggplot(aes(wind, elbo, colour = correct)) + geom_line(aes(group = sim)) +
  labs(x = "Windows", y = "Window ELBO", colour = "Recovery",
       title = "Time = 50") +
  geom_hline(yintercept = -0.2)

elbo_dat %>% 
  filter(time == 1000) %>% 
  filter(wind < 50) %>% 
  mutate(correct = ifelse(clust == 1, "Recover", "Don't")) %>% 
  ggplot(aes(wind, elbo, colour = correct)) + geom_line(aes(group = sim)) +
  geom_hline(yintercept = -0.05) +
  labs(x = "Windows", y = "Window ELBO",  colour = "Recovery",
       title = "Comparison in initial optimisation",
       subtitle = "First 50 windows for T=500") +
  facet_wrap(~correct)

```


## Parameter Recovery

Here we are using 

$$
\frac{1}{K}|\sum B_{ij}-\sum\hat{B}_{ij}|
$$

as the metric for convergence. We see that when the community is
recovered (blue) the matrix is recovered.

```{r param-recovery}
all_b_files <- list.files(path = here("Experiments/exp_results/"),
                          pattern = "exp1_all")

b_ests <- all_b_files %>% 
  map(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y))

sum_b <- b_ests %>% map(~ apply(.x$est_B, 3, sum) )

all_b_df <- b_ests %>% {
  tibble(
    map_dfr(., `[`, c("sim", "clust", "time")),
    sum_B = sum_b
  ) 
} %>% 
  unnest(sum_B) %>% 
  mutate(diff = abs(sum_B - 3.1)/4) %>% 
  group_by(sim) %>% 
  mutate(window = row_number()) %>% 
  ungroup()

all_b_df %>% 
  mutate(Cluster = ifelse(clust == 1, "Recover", "Don't"),
         time = as.factor(time)) %>% 
  # filter(time == 1000) %>% 
  ggplot(aes(window, diff)) + 
  geom_line(aes(group = sim, colour = Cluster)) +
  facet_wrap(~time, scales = "free_x") +
  labs(x = "Observation Period", y = "Difference in Rate",
       subtitle = "78% of Random Initialisation Recover Communities Exactly",
       title = "Recovery of Rate Matrix")

```



# Regret Rate

These experiments make a bit less sense. I'm not sure what to think
of these, maybe a bug here?

Here the regret we are computing is defined as 

$$
Regret(T) = \frac{1}{|A|}
\left(\sum_{m=1}^{M}\tilde{\ell}_{m}(\theta^{(m)}|z^{\ast}) -
    \sum_{m=1}^{M}\tilde{\ell}_{m}(\theta^{\ast}|z^{\ast})\right),
$$

where $\tilde{\ell}$ is the negative log likelihood in a given window.
The $|A|$ is a constant not included in the paper currently.


```{r regret-plot}
exp_3_data <- readRDS(here("Experiments/exp_results/", "exp3_200.RDS"))

exp3_tidy <- exp_3_data %>% tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  dT = list(1:200)
) %>% 
  unnest(cols = c(regret, dT)) %>% 
  select(sim, ARI, regret, dT)

exp3_tidy %>% 
  mutate(ARI = ifelse(ARI == 1, "Recover Community", "Don't Recover")) %>%
  # filter(ARI )
  filter(regret > 0) %>% 
  ggplot(aes(dT, regret)) +
  geom_line(aes(group = sim, colour = ARI), alpha = 0.4) +
  stat_function(fun = function(x) 2*sqrt(x) * log(x*1000)^2,
                colour = "black") +
  facet_wrap(~ARI, scales = "free") +
  labs(y = "Regret Rate", x = "Windows",
       title = "Comparison to theoretical rate",
       subtitle = 
         TeX("Estimated as $2\\sqrt_{T}\\log(|A|T)^2$ with $|A|=1000$.")) +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))

```


# Online Predictive Loss

We can also evaluate the loss for each subsequent window, given 
the current estimates of $\tau$ and $B$. We compare this to
the batch loss, which is computed by fitting the batch estimator 
to the entire dataset and then averaging the predictions for each window from
that model.

We can also show the estimates when we use the true $Z$ vector 
instead of the estimate of $\tau$.

There are scenarios where the predictive loss indicates that
the estimated $Z$ gives a better result than using the true $Z$, but
this is due to label switching which is not incorporated into the
estimates using the true model.


```{r exp4}
exp_4_files <- list.files(path = here("Experiments/exp_results/"), 
                          pattern = "exp4_")

exp4_sims <- exp_4_files %>% 
  map(~readRDS(here("Experiments/exp_results/", .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map_dfr(`[`, c("online_loss", "batch_ave_loss", "sim")) 


exp_4_tidy <- exp4_sims %>% 
  reduce(data.frame) %>% 
  as_tibble() %>% 
  rename(Batch_Loss = elt, Sim = elt.1) %>% 
  group_by(Sim) %>% 
  mutate(Time = max(dT))

```


```{r exp4-50}
curr_time <- 50
exp_4_tidy %>% 
  filter(Time == curr_time) %>% 
  ggplot(aes(dT, Loss, colour = Z)) +
  geom_line() +
  geom_hline(aes(yintercept = Batch_Loss), 
             exp_4_tidy %>% filter(Time == curr_time)) +
  facet_wrap(~Sim, scales = "free")

```

```{r exp4-500}
curr_time <- 500
exp_4_tidy %>% 
  filter(Time == curr_time) %>% 
  ggplot(aes(dT, Loss, colour = Z)) +
  geom_line() +
  geom_hline(aes(yintercept = Batch_Loss), 
             exp_4_tidy %>% filter(Time == curr_time)) +
  facet_wrap(~Sim, scales = "free")
```