---
title: "Thesis Experiments"
author: "Owen G. Ward"
date: "4/18/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path = "rev_figure/",
                      dev = "png",
                      fig.height = 5,
                      fig.width = 7)

library(here)
library(tidyverse)
source(here("functions/utils.R"))

axis_title <- 16
axis_text_size <- 14
legend_text <- 14
line_size <- 1
```

# Figures for the Introduction


## Recovering community structure

```{r fig_1_exp_1}
fig_1_files <- list.files(path = here("Experiments/data_thesis_output/"),
                                      pattern = "fig_1_exp_1")

fig_1_data <- fig_1_files %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))

fig_1_data %>% 
  group_by(Method) %>% 
  summarise(mean(ARI, na.rm = TRUE), sd(ARI, na.rm = TRUE))

# fig_1_data %>% 
#   ggplot(aes(Method, ARI)) +
#   geom_boxplot() +
#   facet_wrap(~Method, scales = "free_x")

method_names <- c("Count", "PZ (window size)", "Point Process")

### want to stratify by window size in PZ also
fig_1 <- fig_1_data %>% 
  mutate(new_x = ifelse(is.na(window_size), " ", window_size)) %>% 
  mutate(new_x = factor(new_x, levels = c(1:10, " "))) %>%
  mutate(Method = if_else(Method == "PZ", "PZ (Window Size)", Method)) %>% 
  mutate(Method = factor(Method,
                         levels = c("Count", "PZ (Window Size)",
                                    "Point Process"))) %>% 
  ggplot(aes(new_x, ARI, fill = Method)) +
  geom_boxplot(alpha = 0.5) +
  facet_wrap(~Method, scales = "free_x") +
  labs(x = element_blank()) +
  theme(legend.position = "none",
        axis.title.y = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        axis.text = element_text(size = axis_text_size),
        legend.text = element_text(size = legend_text))

fig_1
ggsave(here("Thesis_Output", "rev_figures", "fig1.png"),
       fig_1,
       width = 7, height = 5, 
       units = "in", dpi = 300)
  
```



## Interpretable Clusters



# Simulation Experiments

## Community Recovery 


### Increasing the number of nodes

We first wish to confirm that we can recover communities when
data is simulated from this model. Here we show community recovery 
for a fixed $K=2, Time = 200$ setting where we increase the number of
nodes from $n=100,200,500,1000$.
With fixed sparsity $\rho = 0.1$. We do this using a fixed 
$n0$, the proportion of
time used for the initialization, 
and $m0$, the number of nodes used for the initialization.

```{r recov-nodes}
## recovery as the number of nodes changes
node_sims <- list.files(path = here("Experiments", "data_thesis_output"),
                   pattern = "exp_pois_nodes_fixed_may_2")
## to be consistent in the sparsity across all the plots,
## rerunning now

node_data <- node_sims %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))
  

# node_data %>%
#   filter(init == "Init") %>%
#   group_by(nodes, n0, m0) %>%
#   summarise(mean(ARI), sd(ARI))

fig_nodes <- node_data %>% 
  # group_by(sim, nodes, m0) %>% top_n(1) %>% 
  filter(init == "Init") %>%
  # filter(n0 == 10, m0 == nodes/10) %>%
  ggplot(aes(as.factor(nodes), ARI)) +
  geom_boxplot() +
  labs(x = "Number of Nodes") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

ggsave(here("Thesis_Output", "rev_figures", "fig_nodes.png"),
       fig_nodes,
       width = 7, height = 5, 
       units = "in", dpi = 300)

## this looks ok now I guess
```


### Increasing the number of communities


```{r pois_k}

## now just one setting for all, in terms of the sparse initialization 
## scheme

# K_sims <- list.files(path = here("Experiments", "thesis_output"),
#                    pattern = "exp_pois_K_april_25_rho_5")

K_sims <- list.files(path = here("Experiments", "data_thesis_output"),
                   pattern = "exp_pois_K_may_18_rho_5")

K_data <- K_sims %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))

fig_K <- K_data %>% 
  filter(init == "Init") %>%
  filter(nodes == 200) %>% 
  # filter(nodes == 1000) %>% 
  ggplot(aes(as.factor(K), ARI)) +
  geom_boxplot() +
  ylim(c(0, 1)) +
  # facet_wrap(~nodes) +
  labs(x = "K") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

ggsave(here("Thesis_Output", "rev_figures", "fig_k.png"),
       fig_K,
       width = 7, height = 5, 
       units = "in", dpi = 300)
## across multiple community sizes actually looks good too

```

### Online Community Recovery


```{r online_ari}
on_sims <- list.files(path = here("Experiments", "data_thesis_output"),
                   pattern = "exp_pois_online_may_19")


online_dat <- on_sims %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))

online_recov <- online_dat %>% 
  filter(m0 == nodes/4) %>%
  group_by(time, m0, nodes) %>% 
  summarise(Mean_ARI = mean(ARI), sd = sd(ARI)) %>% 
  mutate(nodes = paste0("n = ", nodes)) %>% 
  mutate(nodes = factor(nodes, 
                       levels = c("n = 100", "n = 200", "n = 500",
                                  "n = 1000"))) %>% 
  mutate(lower = max(Mean_ARI - sd, 0), upper = min(1, Mean_ARI + sd)) %>% 
  ggplot(aes(time, Mean_ARI)) +
  geom_point() +
  geom_line(size = line_size) +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  ylim(c(0, 1)) +
  labs(colour = "Number of Nodes", y = "ARI", x = "Time", width = 0.2) +
  facet_wrap(~nodes) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

## need this to be a bit tidier here, but fine

ggsave(here("Thesis_Output", "rev_figures", "online_recov.png"),
       online_recov,
       width = 7, height = 5, 
       units = "in", dpi = 300)
## Old Experiment 9 also has some interesting experiments
```


```{r online-ari-slides, fig.width=7, fig.height=3.5}
online_dat %>% 
  filter(m0 == nodes/4) %>%
  filter(nodes <= 500) %>% 
  group_by(time, m0, nodes) %>% 
  summarise(Mean_ARI = mean(ARI), sd = sd(ARI)) %>% 
  mutate(nodes = paste0("n = ", nodes)) %>% 
  mutate(nodes = factor(nodes, 
                       levels = c("n = 100", "n = 200", "n = 500",
                                  "n = 1000"))) %>% 
  mutate(lower = max(Mean_ARI - sd, 0), upper = min(1, Mean_ARI + sd)) %>% 
  ggplot(aes(time, Mean_ARI)) +
  geom_point() +
  geom_line(size = line_size) +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  ylim(c(0, 1)) +
  labs(colour = "Number of Nodes", y = "ARI", x = "Time", width = 0.2) +
  facet_wrap(~nodes) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))


```


```{r check-this}
online_dat %>% 
  filter(m0 == nodes/4) %>%
  filter(nodes == 100) %>% 
  group_by(time, m0, nodes) %>% 
  summarise(Mean_ARI = mean(ARI), sd = sd(ARI)) %>% 
  mutate(nodes = paste0("n = ", nodes)) %>% 
  mutate(nodes = factor(nodes, 
                       levels = c("n = 100", "n = 200", "n = 500",
                                  "n = 1000"))) %>% 
  mutate(lower = max(Mean_ARI - sd, 0), upper = min(1, Mean_ARI + sd)) %>% 
  ggplot(aes(time, Mean_ARI)) +
  geom_point() +
  geom_line(size = line_size) +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  ylim(c(0, 1)) +
  labs(colour = "Number of Nodes", y = "ARI", x = "Time", width = 0.2) +
  facet_wrap(~nodes) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = axis_text_size),
        legend.title = element_text(size = axis_title))

```

## Monitoring Convergence

Variational inference typically looks at the ELBO for this. 
We can also look at some sort of online version of the ELBO here, 
to evaluate convergence of our online estimation procedure.

```{r monitor_conv_dense}

### look at a single sim here first
### use a loss sim here which stores the ELBO

loss_files <- list.files(path = here("Experiments/data_thesis_output/"), 
                          pattern = "exp_pois_online_loss_april_29")

## how is this normalised? by the cumulative number of events

### try get all of these estimates out at once


elbo_sims <- loss_files %>% 
  map(~read_rds(here("Experiments",
                     "data_thesis_output",
                     .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map_dfr(`[`, c("est_elbo", "sim", "Time")) %>% 
  mutate(est_elbo = as.vector(est_elbo)) %>% 
  group_by(sim) %>% mutate(ind = row_number() + 5)

elbo_sims %>% 
  mutate(Time = paste0("Time = ", Time)) %>% 
  ggplot(aes(ind, est_elbo)) +
  geom_line(aes(group = sim), alpha = 0.2) +
  facet_wrap(~Time, scales = "free") +
  labs(x = "Time", y = "ELBO") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

```

```{r monitor-conv-sparse}
elbo_files <- list.files(path = here("Experiments/data_thesis_output/"), 
                          pattern = "exp_pois_elbo_april")

elbo_sims <- elbo_files %>% 
  map(~read_rds(here("Experiments",
                     "data_thesis_output",
                     .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map_dfr(`[`, c("elbo", "sim", "nodes", "ARI")) %>% 
  mutate(est_elbo = as.vector(elbo)) %>% 
  group_by(sim) %>% mutate(ind = row_number() + 5)

elbo_sims %>% 
  mutate(nodes = paste0("n = ", nodes)) %>% 
  mutate(nodes = factor(nodes, 
                        levels = c("n = 50",
                                   "n = 100",
                                   "n = 200",
                                   "n = 500"))) %>% 
  ggplot(aes(ind, est_elbo)) +
  geom_line(aes(group = sim), alpha = 0.2) +
  # scale_y_continuous(breaks = scales::pretty_breaks(n = 3)) +
  facet_wrap(~nodes) +
  labs(x = "Time", y = "ELBO") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

## sanity check, this looks the same as the community plot
# elbo_sims %>% 
#   group_by(nodes, sim) %>%
#   summarise(ari = mean(ARI)) %>%
#   ungroup() %>% 
#   ggplot(aes(as.factor(nodes), ari)) +
#   geom_boxplot()

### repeat for fixed number of nodes, varying the length of time 
### observed for 


```


```{r elbo-sparse-time}
elbo_time_files <- list.files(path = here("Experiments/data_thesis_output/"), 
                          pattern = "exp_pois_elbo_time")

elbo_time_sims <- elbo_time_files %>% 
  map(~read_rds(here("Experiments",
                     "data_thesis_output",
                     .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map_dfr(`[`, c("elbo", "sim", "nodes", "ARI", "Time")) %>% 
  mutate(est_elbo = as.vector(elbo)) %>% 
  group_by(sim) %>% 
  mutate(ind = row_number() + 5)

elbo_time_sims %>% 
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, levels = c("Time = 50",
                                        "Time = 100",
                                        "Time = 200",
                                        "Time = 500"))) %>% 
  ggplot(aes(ind, est_elbo)) +
  geom_line(aes(group = sim), alpha = 0.2) +
  facet_wrap(~Time, scales = "free") +
  labs(x = "Time", y = "ELBO") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

```


## New ELBO Plot

```{r label, options}

```


## Parameter Recovery


```{r param_recov}
param_data <- loss_files %>% 
     map(~read_rds(here("Experiments",
                         "data_thesis_output",
                         .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map(`[`, c("sim", "Time", "B_ests", "clust")) %>% 
  imap(~update_list(., sum_B = apply(.x$B, 3, sum))) %>% 
  map_dfr(`[`, c("sum_B","sim", "Time", "clust")) %>% 
  group_by(sim, Time) %>% 
  mutate(index = row_number() + 5)

recov_plot <- param_data %>% 
  ### add time here!
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50",
                                  "Time = 100",
                                  "Time = 200",
                                  "Time = 500"))) %>% 
  # filter(clust == 1) %>% 
  mutate(diff = abs(sum_B - 3.1)/4) %>% 
  ggplot(aes(index, diff)) +
  # geom_line(aes(group = sim), alpha = 0.2) +
  geom_smooth(se = FALSE, method = "loess", size = 1) +
  facet_wrap(~Time, scales = "free_x") +
  labs(x = "Time", y = "Difference") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

ggsave(here("Thesis_Output", "rev_figures", "param_recov.png"),
       recov_plot,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```


## Empirical Regret 

```{r exp_regret}
regret_files <- list.files(path = here("Experiments/thesis_output/"),
                          pattern = "exp_pois_regret_april_30")

# regret_data <- regret_sims %>% 
#   map_dfr(~readRDS(here("Experiments/thesis_output/", .x)))

## look at individual sim


sim_50 <- read_rds(here("Experiments",
                             "thesis_output",
                             regret_files[3])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  dT = list(6:50)) %>% 
  unnest(cols = c(regret, dT)) %>% 
  select(sim, ARI, regret, dT) %>% 
  mutate(Time = 50)

sim_100 <- read_rds(here("Experiments",
                             "thesis_output",
                             regret_files[1])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  dT = list(6:100)) %>% 
  unnest(cols = c(regret, dT)) %>% 
  select(sim, ARI, regret, dT) %>% 
  mutate(Time = 100)



sim_200 <- read_rds(here("Experiments",
                             "thesis_output",
                             regret_files[2])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  dT = list(6:200)) %>% 
  unnest(cols = c(regret, dT)) %>% 
  select(sim, ARI, regret, dT) %>% 
  mutate(Time = 200)

sim_500 <- read_rds(here("Experiments",
                             "thesis_output",
                             regret_files[4])) %>%
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  dT = list(6:500)) %>%
  unnest(cols = c(regret, dT)) %>%
  select(sim, ARI, regret, dT) %>%
  mutate(Time = 500)

regret_sims <- bind_rows(sim_50,
                         sim_100,
                         sim_200,
                         sim_500)

# regret_sims %>% 
#   filter(ARI == 1) %>%
#   ggplot(aes(dT, regret)) +
#   geom_line(aes(group = sim), alpha = 0.2) +
#   facet_wrap(~Time, scales = "free") +
#   stat_function(fun = function(x) 0.025 * sqrt(x) * log(x*10000)^2,
#                 colour = "red") +
#   labs(y = "Empirical Regret", x = "Time")


### get smoothed regret
regret_sims %>% 
  filter(ARI == 1) %>%
  group_by(Time, dT) %>% 
  # summarise(q_95 = quantile(regret, 0.95)) %>%
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50", "Time = 100", "Time = 200",
                                  "Time = 500"))) %>% 
  ggplot(aes(dT, regret, colour = Time)) +
  # geom_line(alpha = 0.5) +
  geom_smooth(colour = "black", se = FALSE) +
  facet_wrap(~Time, scales = "free_x") +
  # stat_function(fun = function(x) 0.02 * sqrt(x) * log(x*10000)^2,
  #               colour = "red") +
  labs(y = "Empirical Regret", x = "Time") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))

### maybe without a facet wrap might look better

```


```{r new-regret-plot}
regret_files <- list.files(path = here("Experiments/data_thesis_output/"),
                          pattern = "exp_pois_regret_may_24_")

regret_50 <- list.files(path = here("Experiments/data_thesis_output/"),
                          pattern = "exp_pois_regret_small_may_24_")

regret_files

## then tidy these up, etc
sim_50 <- read_rds(here("Experiments", 
                             "data_thesis_output",
                             regret_files[3])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  perm_regret = map(., "regret_perm"),
  emp_regret = map(., "emp_regret"),
  dT = list(6:50)) %>% 
  unnest(cols = c(regret, perm_regret, emp_regret, dT)) %>% 
  select(sim, ARI, regret, perm_regret, emp_regret, dT) %>% 
  mutate(Time = 50) 

min_perm <- sim_50 %>% 
  group_by(sim) %>% 
  summarise(s1 = sum(regret), s2 = sum(perm_regret)) %>%
  mutate(min = ifelse(s1 < s2, "regret", "perm_regret"))

new_sim_50 <- sim_50 %>% 
  left_join(min_perm %>% select(sim, min), by = "sim") %>% 
  pivot_longer(cols = regret:emp_regret, names_to = "form",
               values_to = "regret") %>% 
  filter((min == form)| form == "emp_regret") %>% 
  mutate(form = ifelse(form == "perm_regret", "regret", form))



## then pivot longer



sim_100 <- read_rds(here("Experiments",
                             "data_thesis_output",
                             regret_files[1])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  emp_regret = map(., "emp_regret"),
  perm_regret = map(., "regret_perm"),
  dT = list(6:100)) %>% 
  unnest(cols = c(regret, perm_regret, emp_regret, dT)) %>% 
  select(sim, ARI, regret, perm_regret, emp_regret, dT) %>% 
  mutate(Time = 100) 


min_perm <- sim_100 %>% 
  group_by(sim) %>% 
  summarise(s1 = sum(regret), s2 = sum(perm_regret)) %>%
  mutate(min = ifelse(s1 < s2, "regret", "perm_regret"))

new_sim_100 <- sim_100 %>% 
  left_join(min_perm %>% select(sim, min), by = "sim") %>% 
  pivot_longer(cols = regret:emp_regret, names_to = "form",
               values_to = "regret") %>% 
  filter((min == form)| form == "emp_regret") %>% 
  mutate(form = ifelse(form == "perm_regret", "regret", form))

# 
# sim_100 %>% 
#   # filter(ARI == 1) %>%
#   group_by(Time, dT, form) %>% 
#   ggplot(aes(dT, regret, colour = form)) +
#   # geom_line(alpha = 0.5) +
#   geom_smooth() 

sim_200 <- read_rds(here("Experiments",
                             "data_thesis_output",
                             regret_files[2])) %>% 
  tibble(
  sim = 1:100,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  perm_regret = map(., "regret_perm"),
  emp_regret = map(., "emp_regret"),
  dT = list(6:200)) %>% 
  unnest(cols = c(regret, perm_regret, emp_regret, dT)) %>% 
  select(sim, ARI, regret, perm_regret, emp_regret, dT) %>% 
  mutate(Time = 200) 


min_perm <- sim_200 %>% 
  group_by(sim) %>% 
  summarise(s1 = sum(regret), s2 = sum(perm_regret)) %>%
  mutate(min = ifelse(s1 < s2, "regret", "perm_regret"))

new_sim_200 <- sim_200 %>% 
  left_join(min_perm %>% select(sim, min), by = "sim") %>% 
  pivot_longer(cols = regret:emp_regret, names_to = "form",
               values_to = "regret") %>% 
  filter((min == form)| form == "emp_regret") %>% 
  mutate(form = ifelse(form == "perm_regret", "regret", form))

sim_500 <- read_rds(here("Experiments",
                             "data_thesis_output",
                             regret_50)) %>%
  tibble(
  sim = 1:50,
  ARI = map_dbl(., "clust"),
  regret = map(., "regret"),
  perm_regret = map(., "regret_perm"),
  emp_regret = map(., "emp_regret"),
  dT = list(6:500)) %>%
  unnest(cols = c(regret, perm_regret, emp_regret, dT)) %>%
  select(sim, ARI, regret, perm_regret, emp_regret, dT) %>%
  mutate(Time = 500)
  
min_perm <- sim_500 %>% 
  group_by(sim) %>% 
  summarise(s1 = sum(regret), s2 = sum(perm_regret)) %>%
  mutate(min = ifelse(s1 < s2, "regret", "perm_regret")) 

new_sim_500 <- sim_500 %>% 
  left_join(min_perm %>% select(sim, min), by = "sim") %>% 
  pivot_longer(cols = regret:emp_regret, names_to = "form",
               values_to = "regret") %>% 
  filter((min == form)| form == "emp_regret") %>% 
  mutate(form = ifelse(form == "perm_regret", "regret", form))


regret_sims <- bind_rows(new_sim_50,
                         new_sim_100,
                         new_sim_200,
                         new_sim_500)

equal_breaks <- function(n = 3, s = 0.05, ...){
  function(x){
    # rescaling
    # d <- s * diff(range(x)) / (1+2*s)
    floor(seq(0, max(x), length=n))
  }
}

dT <- 1
## for the breaks in the facet here

regret_plot <- regret_sims %>% 
  mutate(form = ifelse(form == "regret", "True Membership", form)) %>% 
  mutate(form = ifelse(form == "emp_regret", "Estimated Membership", form)) %>%
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50", "Time = 100", "Time = 200",
                                  "Time = 500"))) %>% 
  filter(ARI == 1) %>%
  filter(form == "Estimated Membership") %>%
  group_by(Time, dT, form) %>% 
  ggplot(aes(dT, regret, colour = form)) +
  # geom_line(alpha = 0.5) +
  geom_smooth(se = FALSE, size = 1) +
  facet_wrap(~Time, scales = "free_x") +
  labs(y = "Regret", x = "Time", colour = "") +
  scale_x_continuous(breaks = equal_breaks(n=3, s=0.05),
                     expand = c(0, max(dT)-1)) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        legend.position = "none")

ggsave(here("Thesis_Output", "rev_figures", "regret_plot.png"),
       regret_plot,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```

## Online Predictions

```{r online-loss}
loss_files <- list.files(path = here("Experiments/data_thesis_output/"), 
                          pattern = "exp_pois_online_loss_april_29")

loss_data <- loss_files %>% 
  map(~readRDS(here("Experiments/data_thesis_output/", .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map_dfr(`[`, c("online_loss", "batch_ave_loss", "sim", "clust")) %>% 
  reduce(data.frame) %>% 
  as_tibble() %>% 
  rename(Batch_Loss = elt, Sim = elt.1, ARI = elt.2) %>% 
  group_by(Sim) %>% 
  mutate(Time = max(dT)) 


batch_loss <- 
  loss_data %>% 
  group_by(Time) %>% 
  summarise(batch = unique(Batch_Loss)) %>% 
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, levels = c("Time = 50",
                                  "Time = 100",
                                  "Time = 200",
                                  "Time = 500")))


loss_plot <- loss_data %>% 
  # filter(Sim == 1) %>% 
  filter(Z == "Est_Z") %>% 
  # filter(Time == 200) %>% 
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50",
                                  "Time = 100",
                                  "Time = 200",
                                  "Time = 500"))) %>% 
  ggplot(aes(dT, Loss)) +
  geom_line(aes(group = interaction(Sim, Z)), alpha = 0.2) +
  geom_hline(data = batch_loss, aes(yintercept =  batch), colour = "red") +
  facet_wrap(~Time, scales = "free_x") +
  labs(x = "Time") +
  theme(legend.position = "null",
        axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),,
        legend.text = element_text(size = legend_text))

ggsave(here("Thesis_Output", "rev_figures", "loss_plot.png"),
       loss_plot,
       width = 7, height = 5, 
       units = "in", dpi = 300)

# single_sim <- loss_data %>% 
#   reduce(data.frame) %>% 
#   as_tibble() %>% 
#   rename(Batch_Loss = elt, Sim = elt.1, ARI = elt.2) %>% 
#   filter(Sim == 1) 
# 
# batch_loss = single_sim$Batch_Loss[1]
# 
# single_sim %>% 
#   ggplot(aes(dT, Loss, colour = Z)) +
#   geom_line() +
#   geom_hline(yintercept = batch_loss)

### then just need to add the batch loss across the bottom

```

```{r loss-plot-slides,fig.width=7, fig.height=3.5}

batch_loss_slides <- batch_loss[1:3, ] 

loss_data %>% 
  filter(Time <= 200) %>% 
  # filter(Sim == 1) %>% 
  filter(Z == "Est_Z") %>% 
  # filter(Time == 200) %>% 
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50",
                                  "Time = 100",
                                  "Time = 200"))) %>% 
  ggplot(aes(dT, Loss)) +
  geom_line(aes(group = interaction(Sim, Z)), alpha = 0.2) +
  geom_hline(data = batch_loss_slides, aes(yintercept =  batch), colour = "red") +
  facet_wrap(~Time, scales = "free_x") +
  labs(x = "Time") +
  theme(legend.position = "null",
        axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title),,
        legend.text = element_text(size = legend_text))


```


# Real Data Experiments

## Event Prediction


College Dataset

```{r college-results}
### get the file path
exps <- list.files(path = here("Experiments/thesis_output"),
                   pattern = "College_Link_pred")

pred_data <- exps %>% 
  map_dfr(~readRDS(here("Experiments/thesis_output/", .x)))

pred_data %>% 
  group_by(model) %>% 
  summarise(median(RMSE), sd(RMSE), median(time), sd(time)) 

## this looks fine, just make it into a nice table

```


```{r email-results}
### get the file path
exps <- list.files(path = here("Experiments/thesis_output"),
                   pattern = "Email_Link_pred")

pred_data <- exps %>% 
  map_dfr(~readRDS(here("Experiments/thesis_output/", .x)))

pred_data %>% 
  group_by(model) %>% 
  mutate(RMSE = ifelse(RMSE > 100, 12.9, RMSE)) %>%
  summarise(median(RMSE), mean(time), sd(time))

### need to figure this one out

```

```{r math-link-pred}

exps <- list.files(path = here("Experiments/thesis_output"),
                   pattern = "Math_Link_pred")

pred_data <- exps %>% 
  map_dfr(~readRDS(here("Experiments/thesis_output/", .x)))

pred_data %>% 
  group_by(model) %>% 
  summarise(median(RMSE), median(time), sd(time)) 

```


```{r ubuntu-link-pred}
exps <- list.files(path = here("Experiments/thesis_output"),
                   pattern = "Ubuntu_Link_pred")

pred_data <- exps %>% 
  map_dfr(~readRDS(here("Experiments/thesis_output/", .x)))

pred_data %>% 
  group_by(model) %>% 
  summarise(median(RMSE), median(time), sd(time)) 

```


# Additional Material

## Varying the window size 

We wish to demonstrate that varying the window size used for
processing the data does not have an appreciable effect on 
the performance of our procedure, in terms of community recovery.


```{r window-size-simulations}
### load in and plot some simulations here

wind_files <- list.files(path = here("Experiments/data_thesis_output"),
                   pattern = "exp_pois_wind_may_23")

wind_data <- wind_files %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))
  

wind_data %>% 
  group_by(dT) %>% 
  summarise(mean(clust))

wind_plot <- wind_data %>% 
  ggplot(aes(as.factor(dT), clust)) +
  geom_boxplot() +
  ylim(c(0, 1)) +
  labs(x = "Window Size", y = "ARI")  +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        strip.text.x = element_text(size = axis_title)) +
  scale_x_discrete(breaks = c(1:5))


ggsave(here("Thesis_Output", "rev_figures", "wind_plot.png"),
       wind_plot,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```


```{r hawkes-nodes}
hawkes_files <- list.files(path = here("Experiments/data_thesis_output"),
                   pattern = "exp_hawkes_new_nodes")


hawkes_sims <- hawkes_files %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))


hawkes_nodes <- hawkes_sims %>% 
  ggplot(aes(as.factor(nodes), ARI)) +
  geom_boxplot() +
  labs(x = "Nodes") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))


ggsave(here("Thesis_Output", "rev_figures", "hawkes_nodes.png"),
       hawkes_nodes,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```


```{r vary-k-hawkes}
hawkes_files <- list.files(path = here("Experiments/data_thesis_output"),
                   pattern = "exp_hawkes_k")


hawkes_sims <- hawkes_files %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))


hawkes_k <- hawkes_sims %>% 
  ggplot(aes(as.factor(K), ARI)) +
  geom_boxplot() +
  labs(x = "K") +
  ylim(c(0, 1)) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))


ggsave(here("Thesis_Output", "rev_figures", "hawkes_k.png"),
       hawkes_k,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```


```{r hawkes-param}

hawkes_par_files <- list.files(path =
                                 here("Experiments/data_thesis_output"),
                   pattern = "exp_hawkes_param")

hawkes_pars <- hawkes_par_files %>% 
  map(~readRDS(here("Experiments/data_thesis_output/", .x))) %>% 
  flatten() %>% 
  imap(~update_list(., sim = .y)) %>% 
  map(`[`, c("sim", "Time", "B_ests", "clust", "Mu_ests", "lambda")) %>% 
  imap(~update_list(., sum_B = apply(.x$B_ests, 3, sum),
                       sum_Mu = apply(.x$Mu_ests, 3, sum))) %>% 
  map_dfr(`[`, c("sim", "Time", "clust", "sum_Mu", "sum_B", "lambda"))


names <- colnames(hawkes_pars)

new_hawkes <- hawkes_pars %>% 
  reduce(data.frame) %>% as_tibble()

colnames(new_hawkes) <- names

hawkes_par <- new_hawkes %>% 
  rename(lam_est = lambda) %>% 
  group_by(sim, Time) %>% 
  mutate(index = row_number()) %>% 
  ### add time here!
  mutate(Time = paste0("Time = ", Time)) %>% 
  mutate(Time = factor(Time, 
                       levels = c("Time = 50",
                                  "Time = 100",
                                  "Time = 200",
                                  "Time = 500"))) %>% 
  mutate(Mu = abs(sum_Mu - 3.65)/4,
         B = abs(sum_B - 1)/4,
         lambda = abs(lam_est - 1)) %>% 
  pivot_longer(cols = Mu:lambda, names_to = "Parameter",
               values_to = "diff") %>% 
  filter(clust == 1) %>% 
  ggplot(aes(index, diff, colour = Parameter)) +
  geom_smooth(se = FALSE, method = "loess", size = line_size) +
  # geom_line(aes(group = interaction(sim, param)), alpha = 0.2) +
  facet_wrap(~Time, scales = "free_x") +
  labs(x = "Time", y = "Difference") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
  scale_fill_discrete(name = "Parameter") +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text))


ggsave(here("Thesis_Output", "rev_figures", "hawkes_par.png"),
       hawkes_par,
       width = 7, height = 5, 
       units = "in", dpi = 300)

```


```{r hawkes-online-community}
### check online commuity recovery

hawkes_files <- list.files(path = here("Experiments/data_thesis_output"),
                   pattern = "exp_hawkes_online")

hawkes_data <- hawkes_files %>% 
  map_dfr(~readRDS(here("Experiments/data_thesis_output/", .x)))

# hawkes_data %>% 
#   group_by(sim) %>% 
#   mutate(final_ari = max(ARI)) %>% 
#   filter(final_ari == 1) %>% 
#   ggplot(aes(time, ARI)) +
#   geom_line(aes(group = sim))


## match above plot
hawkes_online <- hawkes_data %>% 
  group_by(time, nodes) %>% 
  summarise(mean_ARI = mean(ARI), sd_ARI = sd(ARI)) %>% 
  group_by(time) %>% 
  mutate(nodes = paste0("n = ", nodes)) %>% 
  mutate(nodes = factor(nodes, 
                       levels = c("n = 200", "n = 500",
                                  "n = 1000"))) %>% 
  mutate(lower = max(0, mean_ARI - sd_ARI),
         upper = min(1, mean_ARI + sd_ARI)) %>% 
  ggplot(aes(time, mean_ARI)) +
  geom_line(size = 1) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.4) +
  ylim(c(0, 1)) +
  labs(colour = "Number of Nodes", y = "ARI", x = "Time", width = 0.2) +
  facet_wrap(~nodes) +
  theme(axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text)) 


ggsave(here("Thesis_Output", "rev_figures", "hawkes_online.png"),
       hawkes_online,
       width = 7, height = 5, 
       units = "in", dpi = 300)

## do this via colour instead?
## also need to rerun to match the simulation settings for the rest
```



```{r final-exp}
### read in these files and get the summaries


data_path <- list.files(path = here("Experiments/thesis_output"),
                   pattern = "exp_time_")

data_path

sims_data <- data_path %>% 
  map_dfr(~readRDS(here("Experiments/thesis_output/", .x))) %>% 
  mutate(method = c(rep("Poisson", 50),
                    rep("Hawkes", 50),
                    rep("InPois", 50),
                    rep("InHawkes", 50)))


sims_data %>% 
  mutate(time_perc = t_on/t_batch) %>% 
  group_by(method) %>% 
  summarise(median(time_perc), median(perc_diff), mean(perc_diff))


```